---
output: html_document
editor_options: 
  chunk_output_type: console
---

## Distance Between Locations

Determining the distance between locations or between locations and respective habitat types can serve a variety of purposes. Several resource selection procedures require a description of the daily movement distance of an animal to determine the habitat available to an animal or when generating random locations around known locations. We will start here with a method to determine the average distance moved by mule deer in Colorado in a study to determine methods to alleviate depradation on sunflowers that have become a high commodity crop in the area.

1\. Open the script "DistanceUniqueBurst.Rmd" and run code directly from the script

2\. First we need to load the packages needed for the exercise

```{r warning=FALSE,message=FALSE}
library(adehabitatLT)
library(chron)
library(class)
library(sf)
```

3\. Now let's have a separate section of code to include projection information we will use throughout the exercise. In previous versions, these lines of code were within each block of code

```{r warning=FALSE, message=FALSE}
ll.crs <- st_crs(4269)
utm.crs <- st_crs(9001)
albers.crs <- st_crs(5070)
```

4\. Code to read in dataset then subset for an individual animal

```{r}
muleys <-read.csv("data/DCmuleysedited.csv", header=T)
#Code to select an individual animal
muley15 <- subset(muleys, id=="D15")
table(muley15$id)

# #Sort data to address error in code and then look at first 20 records of data to confirm
# muley15 <- muley15[order(muley15$GPSFixTime),]
# #Run code to display the first 20 records to look at what sorting did to data

```

5\. Prepare data to create trajectories using the ltraj command in Adehabitat LT

```{r}
######################################################
## Example of a trajectory of type II (time recorded) with conversion of the date to the 
#format POSIX that nNeeds to be done to get proper digits of date into R then POSIXct uses
#library(chron)
da <- as.character(muley15$GPSFixTime)
da <- as.POSIXct(strptime(muley15$GPSFixTime,format="%Y.%m.%d %H:%M:%S"))
head(da)
#Attach da to muley15
muley15$da <- da

timediff <- diff(muley15$da)
muley15 <-muley15[-1,]
muley15$timediff <-as.numeric(abs(timediff)) 

#Remove outlier locations
coords <- st_as_sf(muley15, coords = c("Long", "Lat"), crs = ll.crs)
plot(st_geometry(coords),axes=T)
deer.spdf <- st_crop(coords, xmin=-110.0,xmax=-106.5,ymin=36.8,ymax=39.0)#Visually identified based on previous plot
plot(st_geometry(deer.spdf),axes=T)

#Project deer.spdf to Albers as in previous exercise
deer.albers <-st_transform(deer.spdf, crs=albers.crs)
plot(st_geometry(deer.albers),axes=T)
```

6\. Create an object of class "ltraj" for muley15 dataset

```{r}
ltraj <- as.ltraj(st_coordinates(deer.albers),deer.albers$da,id=deer.albers$id)
plot(ltraj)
ltraj
#Now let's look at time differences between locations before moving forward
summary(muley15$timediff)

```

7\. Need to create separate "bursts" for each trajectory based on the number of locations collected each day. In our case it was 8 (i.e., locations collected every 3 hours during a 24-hour period).

```{r}
#We want to study the trajectory of the day at the scale of the day. We define one trajectory 
#per day. The trajectory should begin at 2200 hours so the following function returns TRUE if
#the date is time between 06H00 and 23H00 (i.e. results in 7-8 locations/day bursts)
foo <- function(date) {
da <- as.POSIXlt(date)
ho <- da$hour + da$min
return(ho>15.9&ho<23.9)
}
deer <- cutltraj(ltraj, "foo(date)", nextr = TRUE)

#Notice that the above code will remove 328 relocations that fall
#outside of your time criteria
#Warning message:
#In cutltraj(ltraj, "foo(date)", nextr = TRUE) :
#  At least 3 relocations are needed for a burst
# 328 relocations have been deleted
head(deer)
```

8\. Code to change ltraj to a data.frame to summarize distance between locations for each daily burst

```{r eval=FALSE}
dfdeer <- ld(deer)
head(dfdeer)

#Code to get mean distance moved for each burst
dfdeer <- subset(dfdeer, !is.na(dfdeer$dist))#remove NAs from last location of a burst
mean_dist <- do.call(data.frame, aggregate(dfdeer$dist, by=list(dfdeer$burst), 
    function(x) c(mean = mean(x), sd = sd(x), n=abs(length(x)))))
head(mean_dist)
#Write.table gives csv output of Summary 
#write.table(mean_dist, file = "Distance.csv", sep =",", row.names = TRUE, 
#  col.names = TRUE, qmethod ="double")
```
