---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Movement Methods

## Importing datasets from a web source

Movebank.org is a cloud-based repository for relocation data from GPS-collared or VHF-collared animals. It provides a storage facility in the cloud that can serve as a backup for your data or a transfer portal to share data among colleagues or interested researchers. Similar to any email account, each user has a Movebank account that has a login and password to gain access to your data. Administration privileges can be given to anyone with an account for viewing and downloading data.

1\. Open the script "DMAdeerMovebank.Rmd" and run code directly from the script

2\. First we need to load the packages needed for the exercise

```{r warning=FALSE, message=FALSE}
library(move2)
library(RCurl)
library(circular)
library(sf)
```

3\. Next we are going to the [Movebank](https://www.movebank.org/) home page and explore what it has to offer. Need to create an account or select a dataset that does not require permission to use.

```{r eval=FALSE}
movebank_store_credentials("wdwalter")
#login <- movebankLogin(username="wdwalter", password="XXXXX")
```

```{r eval=FALSE}
deer <- movebank_download_study(430427298)
table(deer$individual_local_identifier)

CWD2018.df <- data.frame(st_drop_geometry(deer))
CWD2018.xy <- data.frame(st_coordinates(deer))
CWD2018.new <- cbind(CWD2018.df,CWD2018.xy)

deer2018 <- st_as_sf(CWD2018.new, coords = c("X", "Y"), crs = 4269)
plot(st_geometry(deer2018))

#Now we will select a single deer to explore more
deer1 <- subset(deer2018,deer2018$individual_local_identifier == "12373_12597F")
plot(st_geometry(deer1),add=T, col="red",pch=16)
```
