---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Landscape Metrics

## 7.1 and 7.2 Fragstats Metrics within Polygons

Some research designs may just need landscape metrics for a single area or several study areas and that is what the \\href{<https://r-spatialecology.github.io/landscapemetrics/>}{\emph{landscapemetrics}} package is able to estimate in the code that follows. While a single area can be defined by the extent of the raster imported, the ability of the \emph{landscapemetrics} package to determine patch and class statistics depends on the area defined by the user. This could be a study site, within polygons such as counties or townships, or within buffers around locations.

1\. Open the script "Frag_Auto.Rmd" and run code directly from the script

2\. First we need to load the packages needed for the exercise

```{r warning=FALSE, message=FALSE}
library(plyr)
library(FedData)
#install.packages("landscapemetrics")
library(landscapemetrics)
library(ggplot2)
library(sf)
library(terra)
```

3\. Now let's have a separate section of code to include projection information we will use throughout the exercise. In previous versions, these lines of code were within each block of code

```{r warning=FALSE, message=FALSE}
ll.crs <- st_crs(4269)
utm.crs <- st_crs(9001)
albers.crs <- st_crs(5070)
```

4\. Get land cover raster layer using the FedData pacakage and study area shapefiles

```{r warning=FALSE, message=FALSE, eval=FALSE}
#Load PA shapefile into R
HareCounties <- st_read("data/Hare_Counties.shp")

county_hab <- get_nlcd(template=HareCounties, year = 2006, label = 'HareCounties',dataset = "landcover", force.redo = T)

plot(county_hab)
plot(st_geometry(HareCounties), add=T)

#Let's project Counties to the same projection as the habitat raster
county <- st_transform(HareCounties, albers.crs)
HareCounties <- county
#Matching projections successful!
plot(county_hab)
plot(HareCounties, add=T)

#Now add labels to each hexagon for unique ID
county.centroids <- sf::st_point_on_surface(HareCounties)
county.coords <- as.data.frame(sf::st_coordinates(county.centroids))
county.coords$NAME <- HareCounties$COUNTY_NAM
ggplot() +
  geom_sf(data = HareCounties) +
  geom_text(data = county.coords, aes(X, Y, label = NAME), colour ="black")
```

5\\. Reclassify raster and crop out a single county outline for analysis. Then run landscape metrics on that county

```{r eval=FALSE}
#Grab one county to get landscape metrics
onecounty <- subset(HareCounties, HareCounties$COUNTY_NAM == "FOREST")

# reclassify the values into 3 groups
# all values between 0 and 20 equal 1, etc.
m <- c(-Inf,0,NA, 1, 39, 3, 40, 44, 1, 44.5,52.5,3, 70.5,82.5,2,89.5,95.5,3,95.5, Inf,NA)
rclmat <- matrix(m, ncol=3, byrow=TRUE)
rc <- classify(county_hab, rclmat)
plot(rc)

countyrast <- crop(rc,onecounty, mask=TRUE)
check_landscape(countyrast)

#Calculate a single patch-level metric for euclidian nearest neighbor
ps.data = lsm_p_enn(countyrast)
ps.data

#Calculate the Class statistics for available metrics
cl.data = calculate_lsm(countyrast, level = "class")
cl.data
```

6\. Now we want to export by County name (i.e., COUNTY_NAM) as individual shapefiles. We will only select the first 2 counties for processing to save time

```{r warning=FALSE, message=FALSE, eval=FALSE}
indata <- HareCounties
innames <- unique(HareCounties$COUNTY_NAM)
innames <- innames[1:2]#Place the number of unique polygons in your shapefile here
outnames <- innames

# set up output table
#output <- as.data.frame(matrix(0,nrow=length(innames),ncol=38))

# begin loop to create separate county shapefiles 
for (i in 1:length(innames)){
  data <- indata[which(indata$COUNTY_NAM==innames[i]),]
  if(dim(data)[1] != 0){
    st_write(data, dsn = "data", layer=paste(outnames[i],".shp",sep=""), driver = "ESRI Shapefile")
    write.table(innames, "data/List.txt", col.names=FALSE, quote=FALSE, 
  row.names=FALSE)
}
}

#Read in a list of shapefiles files from above
Listshps<-read.table("data/List.txt",sep="\t",header=F)
Listshps

shape <- function(Listshps) {
file <- as.character(paste("data/", Listshps[1,],".shp", sep=""))
shp <- st_read(file)
mask <- mask(rc,shp)
### Calculate the Class statistics in each county
cl.data <- calculate_lsm(mask, level = "class")
}#or replace patch with "patch" or "landscape" level

results <- ddply(Listshps, 1, shape)
results
#write.table(results, "FragCounty.txt")
```
