---
output: html_document
editor_options: 
  chunk_output_type: console
---

## Creating a Square Polygon Grid Over a Study Area

1\. Open the script "GridSystem2Script.Rmd" and run code directly from the script

2\. First we need to load the packages needed for the exercise

```{r warning=FALSE, message=FALSE}
library(sf)
library(terra)
library(adehabitatMA)
library(FedData)
```

3\. Now let's have a separate section of code to include projection information we will use throughout the exercise. In previous versions, these lines of code were within each block of code

```{r warning=FALSE, message=FALSE}
ll.crs <- st_crs(4269)
utm.crs <- st_crs(9001)
albers.crs <- st_crs(5070)
```

4\. We need to have all layers in same projection so import, create, and remove outliers for mule deer locations then project all to the Albers projection as we did previously.

```{r warning=FALSE, message=FALSE}
#Import the location from earlier exercise
muleys <-read.csv("data/muleysexample.csv", header=T)

#Remove outlier locations
coords <- st_as_sf(muleys, coords = c("Long", "Lat"), crs = ll.crs)
plot(st_geometry(coords),axes=T)
deer.spdf <- st_crop(coords, xmin=-107.0,xmax=-110.5,ymin=37.8,ymax=39.0)#Visually identified based on previous plot
plot(st_geometry(deer.spdf),axes=T)
deer.spdf <-st_transform(deer.spdf, crs=utm.crs)

#Project deer.spdf to Albers as in previous exercise
deer.albers <-st_transform(deer.spdf, crs=albers.crs)
plot(st_geometry(deer.albers,axes=T))
```

5\. Create points for x and y from the bounding box of all mule deer locations with 1500 m spacing between each point.

```{r warning=FALSE, message=FALSE}
bb <- st_bbox(deer.albers) %>% st_as_sfc()
#bb <- cbind(x=c(-108.83966,-108.83966,-108.9834,-108.9834, -108.83966), 
#  y=c(37.8142, 37.86562,37.86562,37.8142,37.8142))
plot(st_geometry(bb))
plot(st_geometry(deer.albers),col="red",add=T)
```

6\. Create a grid of all pairs of coordinates (as a data.frame) using the "expand grid" function and then make it a gridded object.

```{r warning=FALSE, message=FALSE}
grid_spacing <- 500
GridPols <- st_make_grid(bb, square = T, cellsize = c(grid_spacing, grid_spacing)) %>% # the grid, covering bounding box
  st_intersection(bb) %>%
    cbind(data.frame(ID = sprintf(paste("GID%0",nchar(length(.)),"d",sep=""), 1:length(.)))) %>%
    st_sf()
plot(st_geometry(bb))
plot(st_geometry(GridPols),add=T)
```

7\. Similar to the hexagonal grid, identify the cell ID that contains each mule deer location.

```{r}
o = st_intersection(deer.albers,GridPols)
head(o)
```

8\. If we get some NA errors because our grid does not encompass all mule deer locations then we can expand the grid size extending beyond our locations.

```{r}
# Create vectors of the x and y points using boundary box created around deer locations
bb1 <- st_bbox(GridPols)
     
increment = 2000
minx=(min(bb1$xmin)-(increment))
maxx=(max(bb1$xmax)+(increment))
miny=(min(bb1$ymin)-(increment))
maxy=(max(bb1$ymax)+(increment))
  # Create vectors of the x and y points using mean size of deer home range of X square kilometers
x=seq(from=minx,to=maxx,by=increment)
y=seq(from=miny,to=maxy,by=increment)
# Create a grid of all pairs of coordinates (as a data.frame) 
xy=expand.grid(x=x,y=y)
grid.pts<-st_as_sf(xy, coords = c("x","y"),crs = albers.crs)
plot(st_geometry(grid.pts))

GridPols2 <- st_make_grid(grid.pts, square=T,cellsize = grid_spacing) %>%
   cbind(data.frame(ID = sprintf(paste("GID%0",nchar(length(.)),"d",sep=""), 1:length(.)))) %>%
   st_sf()

plot(st_geometry(GridPols2))
plot(st_geometry(GridPols),add=T, col="red")
plot(st_geometry(deer.albers),add=T, col="blue")
```

9.  Then identify the cell ID that contains each mule deer location from the new expanded grid

```{r}
##BE SURE TO RUN CODE FROM XY CREATION THROUGH NEW2 AGAIN THEN LOOK AT DATA!!
o2 = st_intersection(deer.albers,GridPols2)

```

10\. Now we can load a vegetation raster layer using the FedData package to summarize vegetation categories within each polygon grid cell.

```{r}
nlcd <- get_nlcd(template=GridPols2, year = 2019, label = 'nlcd',force.redo = T)
plot(nlcd)
```

11\. Clip the raster within the extent of the newly created grid

```{r}
bbclip <- crop(nlcd, GridPols2)
plot(bbclip)
plot(st_geometry(deer.albers),add=T,col="red")
plot(st_geometry(GridPols2), add=T)

#Cell size of raster layer
xres(bbclip)

#Create histogram of vegetation categories in bbclip
hist(bbclip)

#Calculate cell size in square meters
ii <- st_area(GridPols2)#requires adehabitatMA package
ii[1]
```

12. We can extract the vegetation characteristics within each polygon of the grid and then tabulate area of each vegetation category within each polygon by extracting vegetation within each polygon by ID then summarizing the vegetation characteristics in each cell to be used in future resource selection analysis or disease epidemiology.

```{r}
library(exactextractr)
area = exact_extract(bbclip,GridPols2)
classes <- sort(unique(bbclip[]))
combine <- lapply(area, FUN=function(x) { as.data.frame(prop.table(table(factor(x[,1],
                          levels = classes))))} ) 
head(combine)
```
