---
output: html_document
editor_options: 
  chunk_output_type: console
---

## Creating a Hexagonal Polygon Grid Over a Study Area

1\. Open the script "GridScripts.Rmd" and run code directly from the script

2\. First we need to load the packages needed for the exercise

```{r warning=FALSE, message=FALSE}
library(terra)
library(sf)
library(tigris)
library(FedData)
library(ggplot2)
```

3\. Now let's have a separate section of code to include projection information we will use throughout the exercise. In previous versions, these lines of code were within each block of code

```{r warning=FALSE, message=FALSE}
ll.crs=st_crs(4269)
utm.crs <- st_crs(9001)
albers.crs <- st_crs(5070)
```

4\. We will use the tigris package to downloaded statewide layers for state and county outlines

```{r warning=FALSE, message=FALSE, hide=TRUE}
st <- tigris::states() %>%
  dplyr::filter(GEOID < "60") %>% #GEOID's above 60 are territories and islands, etc. So I'm removing them for scaling.
  tigris::shift_geometry()
plot(st_geometry(st))

st <- st_transform(st, utm.crs)
CO.outline <- subset(st, st$NAME == "Colorado")

COcounties <- counties("Colorado", cb = TRUE)
COcounties <- st_transform(COcounties, utm.crs)
plot(st_geometry(COcounties))

#Import the location from earlier exercise
muleys <-read.csv("data/muleysexample.csv", header=T)

#Remove outlier locations
coords <- st_as_sf(muleys, coords = c("Long", "Lat"), crs = ll.crs)
plot(st_geometry(coords),axes=T)
deer.spdf <- st_crop(coords, xmin=-107.0,xmax=-110.5,ymin=37.8,ymax=39.0)#Visually identified based on previous plot
plot(st_geometry(deer.spdf),axes=T)
deer.spdf <-st_transform(deer.spdf, crs=utm.crs)
```

5\. Now lets extract counties within the extent of the mule deer locations

```{r fig.height=4, fig.width=4, warning=FALSE, message=FALSE}
int <- st_intersection(COcounties,deer.spdf)
clipped <- COcounties[int,]
plot(st_geometry(clipped))
plot(st_geometry(deer.spdf), add=T)
```

6\. We also can create a hexagonal grid across the study site

```{r fig.height=4, fig.width=4, warning=FALSE, message=FALSE}
grid_spacing <- 1500
HexPols <- st_make_grid(clipped, square = F, cellsize = c(grid_spacing, grid_spacing)) %>% # the grid, covering bounding box
  st_intersection(clipped) 
  
plot(st_geometry(clipped))
plot(st_geometry(HexPols),add=T)
```

7\. Create this hexagonal grid across our study site by zooming into deer locations

```{r hide=TRUE, fig.height=4, fig.width=4, warning=FALSE, message=FALSE}
#Import the study site zoomed in shapefile
study.zoom<-st_read("data/MDzoom.shp")
study.zoom <- st_transform(study.zoom, crs=utm.crs)

#Create new hexagonal grid
grid_spacing <- 1500
HexPols <- st_make_grid(study.zoom, square = F, cellsize = c(grid_spacing, grid_spacing)) %>% # the grid, covering bounding box
  st_intersection(study.zoom) %>%
    cbind(data.frame(ID = sprintf(paste("GID%0",nchar(length(.)),"d",sep=""), 1:length(.)))) %>%
    st_sf()
plot(st_geometry(study.zoom))
plot(st_geometry(HexPols),add=T)

#Now add labels to each hexagon for unique ID
hex.centroids <- sf::st_point_on_surface(HexPols)
hex_coords <- as.data.frame(sf::st_coordinates(hex.centroids))
hex_coords$NAME <- HexPols$ID
ggplot() +
  geom_sf(data = HexPols) +
  geom_text(data = hex_coords, aes(X, Y, label = NAME), colour ="black")

```

8\. We can intersect the mule deer locations with the polygon shapefile (i.e., county) they occurred in Hexagon ID if needed

```{r fig.height=4, fig.width=4}
o = st_intersection(deer.spdf,COcounties)

```

9\. As an alternative to importing a polygon that we created in ArcMap above, we can create a polygon in R using the coordinates of the boundary box of the area of interest. In our case here, the bounding box will be the mule deer locations.

```{r fig.height=4, fig.width=4, warning=FALSE, message=FALSE}
bb <- st_bbox(deer.spdf) %>% st_as_sfc()
#bb <- cbind(x=c(-108.83966,-108.83966,-108.9834,-108.9834, -108.83966), 
#  y=c(37.8142, 37.86562,37.86562,37.8142,37.8142))
plot(st_geometry(bb))
plot(st_geometry(deer.spdf),col="red",add=T)
```

10\. Now make practical use of the new bounding box we created by clipping a larger raster dataset. A smaller raster dataset runs analyses faster, provides a zoomed in view of mule deer locations and vegetation, and is just easier to work with.

```{r message=FALSE, eval=FALSE, fig.height=4, fig.width=4}
#Load vegetation raster layer textfile clipped in ArcMap 
veg <-raster("extentnlcd2.txt")
plot(veg)
class(veg)
```

```{r eval=FALSE, warning=FALSE, message=FALSE}
#Clip using the raster imported with "raster" package
bbclip <- st_crop(veg, bb)
veg
```

```{r warning=FALSE, message=FALSE, eval=FALSE}
#WON'T WORK because projections are not the same, WHY?
#Let's check projections of layers we are working with now.
st_crs(MDclip)
st_crs(deer.spdf)
st_crs(SP)
st_crs(veg)
```

11\. We need to have all layers in same projection so project the deer.spdf to Albers and then clip vegetation layer with new polygon we created in the Albers projection.

```{r warning=FALSE, message=FALSE, eval=FALSE}
deer.albers <-st_transform(deer.spdf, crs=albers.crs)
bb.albers <- st_bbox(deer.albers) %>% st_as_sfc()
#Check to see all our layers are now in Albers projection
st_crs(veg)
st_crs(deer.albers)
st_crs(bb.albers)

#Clip using the raster imported with "raster" package 
bbclip <- st_crop(veg, AlbersSP)
plot(bbclip)
plot(st_geometry(deer.albers), col="red", add=T)
plot(st_geometry(bb.albers), lwd=5, add=T)
```
