---
output: html_document
editor_options: 
  chunk_output_type: console
---

## Mid-Atlantic layers using online sources

1\. Open the script "MidAtlantic.Rmd" and run code directly from the script

2\. First we need to load the packages needed for the exercise

```{r warning=FALSE, message=FALSE}
library(sf)
library(terra)
library(FedData)
sessionInfo()#be sure version 3.0 is installed
```

3\. Now let's have a separate section of code to include projection information we will use throughout the exercise

```{r warning=FALSE, message=FALSE}
ll.crs=st_crs(4269)
utm.crs <- st_crs(9001) 
albers.crs <- st_crs(5070) #CRS of shapefile layers #crs <- CRS("+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=23 +lon_0=-96 +x_0=0  #  +y_0=0 +datum=NAD83 +units=m +no_defs +ellps=GRS80 +towgs84=0,0,0") #CRS of raster layers #crs2 <- CRS("+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=23 +lon_0=-96 +x_0=0 +y_0=0  #  +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") 
```

4\. We will use the tigris package to downloaded statewide layers for state and county outlines

```{r}
st <- tigris::states() %>%   dplyr::filter(GEOID < "60") %>%    tigris::shift_geometry() #GEOID's above 60 are territories and islands, etc. So I'm removing them for scaling 
plot(st_geometry(st)) 
st <- st_transform(st, albers.crs) 
PA.outline <- subset(st, st$NAME == "Pennsylvania")  
#PAcounties <- counties("Pennsylvania", cb = TRUE) 
#PAcounties <- st_transform(PAcounties, albers.crs)
#plot(st_geometry(PAcounties))
#We will also get the outline for Maryland to learn to combine rasters
MD.outline <- subset(st, st$NAME == "Maryland")
```

5\. Here we are going to explore a new method to download NLCD or Digital Elevation Data from online within a polygon from your study site.

```{r}
PAnlcd <- get_nlcd(template=PA.outline, label = 'Pennsylvania',force.redo = T)
MDnlcd <- get_nlcd(template=MD.outline, label = 'Maryland',force.redo = T)
nlcd <- mosaic(PAnlcd, MDnlcd, fun=min)
plot(nlcd)
#writeRaster(nlcd,"mosaic_nlcd.tif",format="GTiff",datatype = 'INT4U',overwrite=T)

#nlcd <- rast("mosaic_nlcd.tif")

PADEM <- get_ned(template=PA.outline, label = 'PAdem',force.redo = T)
PADEM.proj <- project(PADEM,PAnlcd,method='near')
#writeRaster(PADEM.proj,"PA_elev.tif",format="GTiff",datatype = 'INT4U',overwrite=T)

#Or read in raster from above
#PADEM.proj <- rast("PA_elev.tif")
plot(PADEM.proj)
plot(st_geometry(PA.outline), add = TRUE)
```

```{r eval=FALSE}
DMA2 <- st_read("data/PGC_CWDManagementUnits2020.shp")
DMA2.proj <- st_transform(DMA2, st_crs(PAnlcd))
DMA_DAYMET <- get_daymet(template=DMA2.proj,label = 'PAdaymet',elements = c('prcp','tmin','tmax'),
                     years = 2010:2011)
plot(DMA_DAYMET$tmin$"2011-05-23")
plot(DMA_DAYMET$tmin$X2011.06.23)
plot(DMA_DAYMET$tmin$X2011.10.23)
plot(DMA_DAYMET$tmin$X2012.05.23)
plot(DMA_DAYMET$prcp$X2011.05.23)

GHCN.prcp <- get_ghcn_daily(template=DMA2, label='GHCNprecip', elements=c('prcp'))
plot(st_geometry(DMA2))
plot(st_geometry(GHCN.prcp$spatial), add=T,pch=1)
legend('topright', pch=1, legend='GHCN Precipitation Records')
```

```{r eval=FALSE}
DMAsoil <- get_ssurgo(template=PAstate, label = 'DMA_SSURGO',force.redo = T) 


```

```{r eval=FALSE}
NHD <- get_nhd(template=PAstate, label='PAnhd')%>%
  plot_nhd(template = PAstate)
get_nhd(template=PAstate, label='PAnhd')%>%
  plot_nhd(template = PAstate)
plot(DMA2.proj)
plot(NHD$NHDFlowline, add=T)
plot(NHD$NHDLine, add=T)
plot(NHD$NHDArea, col='black', add=T)
plot(NHD$NHDWaterbody, col='black', add=T)
```
