---
output: html_document
editor_options: 
  chunk_output_type: console
---

## Movement Trajectories

We will start with simply creating trajectories between successive locations. As stated previously, there are 2 types of trajectories but their are also 2 forms of Type II trajectories if we have time recorded. Depending on the duration between locations we can have uniform time lag between successive relocations termed regular trajectories and non-uniform time lag that results in irregular trajectories. We will begin this section with simply creating irregular trajectories from relocation data because, even though we set up a time schedule to collection locations at uniform times, climate, habitat, and satellites do not always permit such schedules of data collection.

1\. Open the script "MovementScript.Rmd" and run code directly from the script

2\. First we need to load the packages needed for the exercise

```{r warning=FALSE, message=FALSE}
library(adehabitatLT)
library(chron)
library(sf)
```

3\. Now let's have a separate section of code to include projection information we will use throughout the exercise. In previous versions, these lines of code were within each block of code

```{r warning=FALSE, message=FALSE}
ll.crs <- st_crs(4269)
utm.crs <- st_crs(9001)
albers.crs <- st_crs(5070)
```

4\. We are again going to be using more of the mule deer dataset than from the earlier exercises

```{r}
muleys <-read.csv("data/DCmuleysedited.csv", header=T)
```

5\. Check for duplicate locations in dataset. The reason for this is very important and will be apparent shortly.

```{r fig.height=4, fig.width=4}
#Check for duplicate locations in dataset
summary(duplicated(muleys))
```

6\. For trajectories of type II (time recorded), the conversion of the date to the format POSIX needs to be done to get proper digits of date into R.

```{r}
da <- as.POSIXct(strptime(muleys$GPSFixTime,format="%Y.%m.%d %H:%M:%S"))
muleys$da <- da

timediff <- diff(muleys$da)*60
muleys <-muleys[-1,]
muleys$timediff <-as.numeric(abs(timediff)) 

#Remove outlier locations
coords <- st_as_sf(muleys, coords = c("Long", "Lat"), crs = ll.crs)
plot(st_geometry(coords),axes=T)
deer.sf <- st_crop(coords, xmin=-107.0,xmax=-110.5,ymin=37.8,ymax=39.0)#Visually identified based on previous plot
plot(st_geometry(deer.sf),axes=T)

#Project deer.spdf to Albers as in previous exercise
deer.albers <-st_transform(deer.sf, crs=albers.crs)
plot(st_geometry(deer.albers),axes=T)

```

6\. Now create an object of class "ltraj" by animal using the ID field and display by each individual (i.e., ltraj[1])

```{r}
ltraj <- as.ltraj(st_coordinates(deer.albers),deer.albers$da,id=deer.albers$id)
head(ltraj[1])#Describes the trajectory
plot(ltraj)#plot all trajectories created
#Plot each trajectory separately
plot(ltraj[1])
plot(ltraj[2])
plot(ltraj[3])
plot(ltraj[4])
plot(ltraj[5])
plot(ltraj[6])
```

7\. Create a histogram of time lag (i.e., interval) and distance between successive locations for each deer. This is a nice way to inspect the time lag between locations as you don't want to include a location if too much time has passed since the previous and it also shows why a trajectory is irregular.

```{r eval=FALSE}
hist(ltraj[1], "dt", freq = TRUE)
hist(ltraj[1], "dist", freq = TRUE)
hist(ltraj[2], "dt", freq = TRUE)
hist(ltraj[2], "dist", freq = TRUE)
hist(ltraj[3], "dt", freq = TRUE)
hist(ltraj[3], "dist", freq = TRUE)
hist(ltraj[4], "dt", freq = TRUE)
hist(ltraj[4], "dist", freq = TRUE)
hist(ltraj[5], "dt", freq = TRUE)
hist(ltraj[5], "dist", freq = TRUE)
hist(ltraj[6], "dt", freq = TRUE)
hist(ltraj[6], "dist", freq = TRUE)
```
