---
output: html_document
editor_options: 
  chunk_output_type: console
---

## Kernel density estimators (KDE) with various bandwidth selection using package amt

Here we will estimate home range using several kernel density estimators on the same animal in the package amt. We will conclude with estimating home range using Brownian Bridge Movement Models (BBMM) and autocorrelated kernel density estimator developed for autocorrelated global positioning system datasets for comparison to traditional kernel density estimators.

1\. Open the script "Panther_All4_amt.Rmd" and run code directly from the script

2\. First we need to load the packages needed for the exercise

```{r warning=FALSE, message=FALSE}
library(adehabitatHR)
library(amt)
library(stringr)
library(PBSmapping)
library(sf)
library(terra)
library(wildlifeDI)#for sf2ltraj function in new BBMM code
library(ggplot2)
library(units)
library(move2)
```

3\. Now let's have a separate section of code to include projection information we will use throughout the exercise. In previous versions, these lines of code were within each block of code

```{r warning=FALSE, message=FALSE}
ll.crs <- st_crs(4269)
utm.crs <- st_crs(26917)
albers.crs <- st_crs(5070)
```

4\. We will use an abbreviated dataset to save processing time and the code will also output shapefiles of home ranges

```{r}
panther<-read.csv("data/pantherjitter.csv",header=T)
panther$CatID <- as.factor(panther$CatID)
panther$NewTime <- str_pad(panther$TIMEET2,4, pad= "0")
panther$NewDate <- paste(panther$DateET2,panther$NewTime)
#Used to sort data in code below for all deer
panther$DT <- as.POSIXct(strptime(panther$NewDate, format='%Y %m %d %H%M'))
#Sort Data
panther <- panther[order(panther$CatID, panther$DT),]
#TIME DIFF NECESSARY IN BBMM CODE
timediff <- diff(panther$DT)*60
# remove first entry without any difference 
panther <- panther[-1,] 
panther$timelag <-as.numeric(abs(timediff))

cat143 <- subset(panther, panther$CatID == "143")
cat143$CatID <- droplevels(cat143$CatID)

cat143track <- make_track(tbl=cat143, .x=X, .y=Y, .t=DT, id=CatID, crs=26917)

```

5\. We will start by running KDE with href similar to previous href exercise.

```{r warning=FALSE, message=FALSE}
## If we want to see what bandwidth the model will be using, we can use the following
hr_kde_ref(cat143track)

## To estimate the kde, we'll use the hr_kde function and specify that h is the reference bandwidth (h=hr_kde_ref). For this exercise, we are going to calculate 50%, 80%, 90%, 95% and 99% home ranges, so we'll specify those using the levels argument (although you can calculate any level of home range)

cat143.href <- hr_kde(cat143track, h=hr_kde_ref(cat143track), levels=c(0.5,0.80,0.90,0.95,0.99))

image(cat143.href$ud)

## The heatmap above isn't really useful - what we are really interested in are the home range isopleths (i.e. the boundaries of our different home range percentages) and the sizes of those boundaries.
cat143.href.iso<-hr_isopleths(cat143.href, levels=c(0.5,0.80,0.90,0.95,0.99))

## The previous line of code creates an sf dataframe which included the level, the area estimate, and the geometry (to be used for plotting). The default area measurement is m^2, but we are going to change it to km2 to make it easier to interpret
cat143.href.iso$area<-set_units(cat143.href.iso$area, km^2)

g1 <- ggplot(cat143.href.iso)+
  geom_sf(aes(fill=as.factor(level)))+
  scale_fill_manual(name="Isopleth", values=c("blue4","royalblue2","steelblue2", "turquoise2","yellow1"), labels=c("50%", "80", "90", "95%", "99%"))+ ## This addes a legend title, changes the colors of the isopleths and changes the labels
  theme_bw()+
  theme(panel.grid = element_blank(), axis.text = element_blank())+
  ggtitle("KDE-HREF Bandwidth")
g1
```

6\. Next we will run KDE with hplug-in similar to previous hplug-in exercise.

```{r  warning=FALSE, message=FALSE}
## We will use the same elk track that we used before and can see what h values are estimated with the plug-in approach
hr_kde_pi(cat143track)

## The function to estimate the home range is the same as for href, except we change the h argument to hr_kde_pi().
cat143.plugin<-hr_kde(cat143track, h=hr_kde_pi(cat143track), levels=c(0.5,0.80,0.90,0.95,0.99))
image(cat143.plugin$ud)

cat143.plugin.iso<-hr_isopleths(cat143.plugin, levels=c(0.5,0.80,0.90,0.95,0.99))

cat143.plugin.iso$area<-set_units(cat143.plugin.iso$area, km^2)

g2 <- ggplot(cat143.plugin.iso)+
  geom_sf(aes(fill=as.factor(level)))+
  scale_fill_manual(name="Isopleth", values=c("blue4","royalblue2","steelblue2", "turquoise2","yellow1"), labels=c("50%", "80", "90", "95%", "99%"))+ ## This addes a legend title, changes the colors of the isopleths and changes the labels
  theme_bw()+
  theme(panel.grid = element_blank(), axis.text = element_blank())+
  ggtitle("KDE-Plug-in Bandwidth")
g2
```

7\. Next we will run BBMM. Because BBMM package is no longer maintained on CRAN, we will use the function provided by adehabitatHR package with plotting adapted to sf and ggplot2 packages

```{r}
#loc <- st_as_sf(x = cat143, coords = c("X","Y"),crs = utm.crs)
loc <- mt_as_move2(cat143, coords = c("X", "Y"), na.fail = FALSE, time_column = "DT", track_id_column = "CatID",crs = utm.crs)

cat143traj <- move2_ltraj(loc)

sig_df<-(liker(cat143traj, sig2 = 10, rangesig1 = c(1, 100)))

sig1_value<-sig_df[[1]]$sig1

BBMM = kernelbb(cat143traj, sig1 = sig1_value, sig2 = 10, grid = 100)

image(BBMM)

hr99<-getverticeshr(BBMM, 99, unout = "km2")
hr99
hr99_sf<-st_as_sf(hr99)

hr95<-getverticeshr(BBMM, 95, unout="km2")
hr95
hr95_sf<-st_as_sf(hr95)

hr90<-getverticeshr(BBMM, 90, unout="km2")
hr90
hr90_sf<-st_as_sf(hr90)

hr80<-getverticeshr(BBMM, 80, unout="km2")
hr80
hr80_sf<-st_as_sf(hr80)

hr50<-getverticeshr(BBMM, 50, unout="km2")
hr50
hr50_sf<-st_as_sf(hr50)

g3 <- ggplot(hr99_sf)+
  geom_sf(aes(fill="99%"))+
  geom_sf(data=hr95_sf, aes(fill="95%"))+ 
  geom_sf(data=hr90_sf, aes(fill="90%"))+
  geom_sf(data=hr80_sf, aes(fill="80%"))+
  geom_sf(data=hr50_sf, aes(fill="50%"))+
  scale_fill_manual(name="Isopleth", values=c("blue4","royalblue2","steelblue2", "turquoise2","yellow1"))+
  theme_bw()+
  theme(panel.grid = element_blank(), axis.text = element_blank())+
  ggtitle("BBMM")
g3
```

8. We will quickly explore autocorrelated kernel density estimator which uses the "auto" option to choose the best movement model to estimate home range.

```{r}
cat143.akde <- hr_akde(cat143track, model=fit_ctmm(cat143track, "iid"), levels=c(0.5,0.80,0.90,0.95,0.99))

cat143.akde.iso<-hr_isopleths(cat143.akde, levels=c(0.5,0.80,0.90,0.95,0.99))

plot(cat143.akde)

cat143.akde.iso$area<-set_units(cat143.akde.iso$area, km^2)

g4 <- ggplot(cat143.akde.iso)+
  geom_sf(aes(fill=as.factor(level)))+
  scale_fill_manual(name="Isopleth", values=c("blue4","royalblue2","steelblue2", "turquoise2","yellow1"), labels=c("50%", "80", "90", "95%", "99%"))+ ## This addes a legend title, changes the colors of the isopleths and changes the labels
  theme_bw()+
  theme(panel.grid = element_blank(), axis.text = element_blank())+
  ggtitle("AKDE")
g4
```

9. We can add 4 estimators to the plot window to compare across estimators using the patchwork package

```{r}
library(patchwork)
g1+g2+g3+g4+plot_layout(ncol=2)
```

```{r}
## We also want to make a plot showing how the area differs by estimator. We'll first add labels to the href and hpi dataframes
cat143.href.iso$estimator<-"H Ref"
cat143.plugin.iso$estimator<-"H Plug-in"

## The akde dataframe also includes confidence interval values for the area estimates. We aren't going to include those for this plot, so we will filter out only the estimate values, and also add a label column
cat143.akde.iso2<-filter(cat143.akde.iso, what=="estimate")
cat143.akde.iso2$estimator<-"AKDE"

## The bbmm output was stored differently, so we need to make a dataframe that matches we we have for the other estimators. We'll first use the rbind function to store the area associated with the 3 area measurements
bbmm.area<-as.data.frame(rbind(hr50$area, hr80$area, hr90$area, hr95$area, hr99$area))

## Name the column the same as it is in the other estimators' dataframes ("area")
colnames(bbmm.area)<-"area"

## Add a new column that specifies the level of each measurement, and a label column 
bbmm.area<-bbmm.area%>%mutate(level=c(0.50,0.80,0.90,0.95,0.99))
bbmm.area$estimator<-"bbmm"

## Bind together the href, hpi, and akde dataframes, drop the geometry and units from the resulting dataframe
hr.area<-rbind(cat143.href.iso,cat143.plugin.iso, cat143.akde.iso2)
hr.area<-st_drop_geometry(hr.area)
hr.area$area<-drop_units(hr.area$area)

## Rearrange the columns so that they match what is in the bbmm data frame(area, level, estimator)
hr.area<-hr.area[,c(3,1,4)]

## Bind all the estimators' dataframes together and store level as a factor
hr.area.all<-rbind(hr.area, bbmm.area)
hr.area.all$level<-as.factor(hr.area.all$level)

## For graphing purposes, we want to order the estimators from highest estimate to lowest estimate. There are better ways to do this, but for the sake of time we are going to look at the values and manually re-order them
hr.area.all$estimator<-factor(hr.area.all$estimator, levels=c("AKDE", "H Ref", "bbmm", "H Plug-in"))


ggplot(hr.area.all, aes(y=area, x=level, color=estimator, fill=estimator))+
  geom_bar(stat="identity",position=position_dodge())+
  scale_x_discrete(labels=c("50%","80%","90%", "95%", "99%"))+
  scale_color_discrete(name="Estimator")+
  scale_fill_discrete(name="Estimator")+
  scale_y_continuous(expand=c(0,0))+ ## This forces the bar to start at 0. ggplot normally has a space before 0 which I find annoying
  xlab("Home range isopleth")+
  ylab(bquote("Average home range size  " (km^2)))+
  theme_classic()
```
