---
output: html_document
editor_options: 
  chunk_output_type: console
---

## Manipulate Raster Data Layer

1\. Open the script "RasterScript.R"

2\. First we need to load the packages needed for the exercise

```{r warning=FALSE, message=FALSE}
#install.packages(c("adehabitatHR","maptools","raster", "rgdal"))
library(terra)
library(sf)
library(dplyr)
library(tigris)
#options(tigris_use_cache = TRUE)
#tigris_cache_dir("/Users/davidwalter/Library/CloudStorage/OneDrive-ThePennsylvaniaStateUniversity/WalterRprojects/Manual-Applied-Spatial-Ecology/data")
#'
#install.packages("remotes")
# remotes::install_github(repo = "r-lib/devtools",
#                               dependencies = TRUE,
#                               upgrade = TRUE)
# devtools::install_github(repo = "r-lib/devtools",
#                            dependencies = TRUE,
#                            upgrade = TRUE)
#devtools::install_github("ropensci/FedData")
#remotes::install_github("ropensci/FedData")
library(FedData)
sessionInfo() #Search for and confirm FedData_4.0.0 or newer was loaded
#library(adehabitatHR)
#library(maptools)
```

3\. Now let's have a separate section of code to include projection information we will use throughout the exercise

```{r warning=FALSE, message=FALSE}
ll.crs=st_crs(4269)
utm.crs <- st_crs(32617)
albers.crs <- st_crs(5070)
#CRS of shapefile layers
#crs <- CRS("+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=23 +lon_0=-96 +x_0=0 
#  +y_0=0 +datum=NAD83 +units=m +no_defs +ellps=GRS80 +towgs84=0,0,0")
#CRS of raster layers
#crs2 <- CRS("+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=23 +lon_0=-96 +x_0=0 +y_0=0 
#  +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs")

```

4\. We will use the tigris package to downloaded statewide layers for state and county outlines

```{r eval=FALSE}
st <- tigris::states() %>%
  dplyr::filter(GEOID < "60") %>% 
  tigris::shift_geometry()
#GEOID's above 60 are territories and islands, etc. So I'm removing them for scaling.
plot(st_geometry(st))

st <- st_transform(st, albers.crs)
SC.outline <- subset(st, st$NAME == "South Carolina")

SCcounties <- counties("South Carolina", cb = TRUE)
SCcounties <- st_transform(SCcounties, albers.crs)
```

5\. Now let's add some shapefiles to our raster specific to the military base of interest

```{r  eval=FALSE, warning=FALSE, message=FALSE, hide=TRUE}
#Load county shapefile
county<- subset(SCcounties, SCcounties$NAME == "Beaufort")

#Load airport runway shapefile
run<-st_read("data/RunwayAlbers.shp")

#Load aircraft flight pattern shapefile
path<-st_read("data/FlightImage.shp")

#Load aircraft flight pattern shapefile
road <- st_read("data/CountyRoadAlbers.shp")
```

6.  Using county outline, we can then download National Land Cover data for any year of interest using the FedData package

```{r warning=FALSE, message=FALSE, eval=FALSE}
#new.outline <- sf::as_Spatial(county)#need to do this line due to error with dplyr package
SCnlcd <- get_nlcd(template=SC.outline, year = 2019, label = 'SC',force.redo = T)
#Write raster if needed
#writeRaster(SCnlcd, filename = "SCnlcd2019.tif", filetype = "GTiff", datatype = 'INT4U',overwrite=TRUE)
  
plot(SCnlcd)
plot(st_geometry(SCcounties), add=T, lwd=3)

#Import raster saved above just for demonstration
#SCnlcd <-rast("SCnlcd2019.tif")
plot(SCnlcd)
plot(st_geometry(SCcounties), add=T, lwd=3)
```

7\. Plot out all the shapefiles overlayed on each other with and without the raster.

```{r eval=FALSE}
plot(st_geometry(county))
plot(st_geometry(road), add=T)
plot(st_geometry(run), col="red",add=T)
plot(st_geometry(path), col="blue",add=T)
```

8\. Let's reclassify NLCD layer to get fewer land cover categories to make it easier to work with the raster.

```{r eval=FALSE}
# reclassify the values into 7 groups all values between 0 and 20 equal 1, etc. while removing water (11)
m <- c(0, 19, NA, 20, 39, 1, 40, 50, 2, 51, 68, 3, 69,79, 4, 80, 88, 5, 89, 99, 6)
rclmat <- matrix(m, ncol=3, byrow=TRUE)
rc <- classify(SCnlcd, rclmat)
plot(rc)

#Clip the raster within the county polygon for a zoomed in view then plot
clip <- crop(rc, county)
plot(clip)
plot(st_geometry(county),add=T,lwd=2)
plot(st_geometry(run),add=T,lwd=2, col="red")
plot(st_geometry(path),add=T,lty="62", col="blue")
plot(st_geometry(road),add=T,lty="22", col="yellow")
```

9\. We can load some vulture locations to extract landcover that each location occurs in that will be considered "used" habitat in resource selection analysis.

```{r eval=FALSE}
#Import bird 49 locations to R
bv49 <-read.csv("data/Bird49.csv", header=T)#How many bird locations?

bvsf <- st_as_sf(bv49, coords = c("x", "y"), crs = utm.crs)

plot(st_geometry(bvsf), col="red")
#NOTE: Locations must be assigned to the UTM coordinate system prior to projection to Albers 
#so won't overlay on veg layer at this point because veg is in Albers
bv49Albers <- st_transform(bvsf, crs = albers.crs)
plot(clip)
plot(st_geometry(bv49Albers), col="blue",add=T)
```

10\. Determine which of those points lie within a cell that contains data by using the extract function. The extract function will extract covariate information from the raster for xy coordinates at for each location.

```{r eval=FALSE}
bvspdf.df <- data.frame(st_coordinates(bv49Albers))
bvspdf2 <- st_as_sf(bvspdf.df, coords= c("X", "Y"), crs = albers.crs)
bvspdf.sv <- vect(bvspdf2)#Need SpatVector for extract function in terra
veg.survey<-terra::extract(clip, bvspdf.sv)
veg.survey<-subset(veg.survey,!is.na(veg.survey$Class))
```

11\. We can also create some random points within the extent of the area to be considered as "available" habitat.

```{r eval=FALSE}
##First we need to create a box across the study site to generate a random sample of locations
e.box <- as.polygons(clip,extent=TRUE)
e.box <- st_as_sf(e.box)
Sample.points <- st_sample(e.box, 1000, type="random")

plot(clip)
plot(st_geometry(Sample.points),add=T)

#Determine which of those points lie within a cell that contains data by using the extract function again.
Sample.points.sv <- vect(Sample.points)#Need SpatVector class for extract function in terra
samp.survey<-terra::extract(clip, Sample.points.sv)
samp.survey<-subset(samp.survey,!is.na(samp.survey$Class))
```

12\. New code addition below (Update 1/7/2021) to use mapview package to treat plotting data layers as done with GIS software. This package enables user to zoom in and out and scroll around layer after calling a plot function \newpage

```{r eval=FALSE}
library(mapview)
library(stars)
mapview(st_as_stars(clip)) + mapview(run) + mapview(path)
mapview(county) + mapview(bv49Albers, color="yellow")
mapview(run) + bv49Albers + road
```
