## Regular Trajectories

1\. Open the script "RegTrajScript.Rmd" and run code directly from the script

2\. First we need to load the packages needed for the exercise

```{r warning=FALSE,message=FALSE}
library(adehabitatLT)
library(chron)
library(sf)
```

3\. Now let's have a separate section of code to include projection information we will use throughout the exercise. In previous versions, these lines of code were within each block of code

```{r warning=FALSE, message=FALSE}
ll.crs <- st_crs(4269)
utm.crs <- st_crs(9001)
albers.crs <- st_crs(5070)
```

4\. Now read in dataset, extract a single animal and create ltraj as in previous exercise

```{r fig.height=4, fig.width=4}
muleys <-read.csv("data/DCmuleysedited.csv", header=T)

#CODE FOR AN INDIVIDUAL ANIMAL
muley15 <- subset(muleys, id=="D15")
muley15$id <- factor(muley15$id)
table(muley15$id)

#Sort data to address error in code and then look at first 10 records of data to confirm
muley15 <- muley15[order(muley15$GPSFixTime),]

######################################################
## Example of a trajectory of type II (time recorded)
### Conversion of the date to the format POSIX
#Needs to be done to get proper digits of date into R then POSIXct
#uses library(chron)
da <- as.character(muley15$GPSFixTime)
da <- as.POSIXct(strptime(muley15$GPSFixTime,format="%Y.%m.%d %H:%M:%S"))
#Attach da to muley15
muley15$da <- da

timediff <- diff(muley15$da)
muley15 <-muley15[-1,]
muley15$timediff <-as.numeric(abs(timediff)) 

#Remove outlier locations
coords <- st_as_sf(muley15, coords = c("Long", "Lat"), crs = ll.crs)
plot(st_geometry(coords),axes=T)
deer.spdf <- st_crop(coords, xmin=-110.0,xmax=-106.5,ymin=36.8,ymax=39.0)#Visually identified based on previous plot
plot(st_geometry(deer.spdf),axes=T)

#Project deer.spdf to Albers as in previous exercise
deer.albers <-st_transform(deer.spdf, crs=albers.crs)
plot(st_geometry(deer.albers),axes=T)

#Creation of an object of class "ltraj"
ltraj <- as.ltraj(st_coordinates(deer.albers),deer.albers$da,id=deer.albers$id)
plot(ltraj)
```

6.We want to study the trajectory of the day at the scale of the day. We define one trajectory per day. The trajectory should begin at 2200 hours so the following function returns TRUE if the date is time between 06H00 and 23H00 (i.e. results in 7-8 locations/day bursts)

```{r fig.height=4, fig.width=4}
foo <- function(date) {
da <- as.POSIXlt(date)
ho <- da$hour + da$min
return(ho>18.0&ho<23.9)
}
deer <- cutltraj(ltraj, "foo(date)", nextr = TRUE)
head(deer)

## Remove the first and last burst if needed?
#deer2 <- deer[-c(1,length(deer))]

#Bind the trajectories
deer3 <- bindltraj(deer)
deer3
plot(deer3)
is.regular(deer3)
plotltr(deer3, "dt")

## The relocations have been collected every 3 hours, and there are some
## missing data
## The reference date: the hour should be exact (i.e. minutes=0):
refda <- strptime("00:00", "%H:%M")
refda
## Set the missing values
deerset <- setNA(deer3, refda, 3, units = "hour")
## now, look at dt for the bursts:
plotltr(deerset, "dt")
## dt is nearly regular: round the date:
deerset1 <- sett0(deerset, refda, 3, units = "hour")
plotltr(deerset1, "dt/3600")
is.regular(deerset1)
## deerset1 is now regular

## Is the resulting object "sd" ?
is.sd(deerset1)


#Show the changes in the distance between successive relocations with the time
plotltr(deerset1, "dist")
deerset1#Is the trajectory regular now?
```
