---
output: html_document
editor_options: 
  chunk_output_type: console
---

## Manipulate Polygon Layer

1\. Open the script "SoilScript.Rmd" and run code directly from the script

2\. First we need to load the packages needed for the exercise

```{r warning=FALSE, message=FALSE}
library(sf)
library(terra)
```

3\. Import shapefile of soils dataset

```{r , message=FALSE, warning=FALSE}
soils<-st_read("data/Soil_Properties.shp")
plot(st_geometry(soils))
names(soils) #get attribute data

#Rename original category headings to something more familiar
soils$Clay <- soils$SdvOutpu_1
soils$pH <- soils$SdvOutpu_2
soils$CEC <- soils$SdvOutpu_3
```

4.Shapefiles contain several features associated with each polygons that collectively make up the entire layer so let's explore

```{r eval=FALSE}
soils #a data frame with geometry type "polygon" associated with 10 features
st_bbox(soils) #boundary box
st_crs(soils) #projection information
soils[1, ] #will bring up data associated with the first polygon
```

5\. Select portions of the data that fit some set criteria

```{r}
#Highlights the areas that Percent Clay polygons are over 30%
plot(st_geometry(soils))
high.clay <- soils[soils$Clay>30,]
plot(st_geometry(high.clay), border="red", add=TRUE)

##Highlights the areas that Cation Exchange Capacity is greater than 14
high.CEC<- soils[soils$CEC>14,]
plot(st_geometry(high.CEC), border="green", add=TRUE)

##Highlights the areas that soil pH is greater than 8
high.pH <- soils[soils$pH>8,]
plot(st_geometry(high.pH), border="yellow", add=TRUE)
```

6\. Bring in locations of harvested mule deer

```{r}
#Import mule deer locations from harvested animals tested for CWD
#Note: Locations have been offset or altered so do not reflect actual locations of samples
mule <- read.csv("data/MDjitterclip.csv", header=T)
crs<-"+proj=utm +zone=13 +datum=WGS84 +no_defs +towgs84=0,0,0"
coords <- st_as_sf(mule, coords = c("x", "y"), crs = st_crs(soils))

plot(st_geometry(coords), col="blue")
#par(new=TRUE)
```

7\. Let's generate random points with the extent of the soil layer

```{r}
#Sampling points in a Spatial Object "type=regular" will give a regular grid
samples<-st_sample(soils, 1000, type="random")

plot(st_geometry(soils), col="wheat")
plot(st_geometry(coords), col="blue", add=T)
plot(st_geometry(samples), col="red", add=T)
```

8\. Extract and tally Clay soil types for random samples and mule deer locations

```{r}
#Matches points with polygons:
soils.idx<- st_intersects(samples,soils)
soil.samples <- soils[unlist(soils.idx), "Clay", drop = TRUE] # drop geometry
# locs <- SpatialPoints(coords)
# locs@proj4string <- soils@proj4string
soils.locations<- st_intersects(coords, soils)
soils.locs <- soils[unlist(soils.locations), "Clay", drop = TRUE]
#Tally clay soil types for random samples
obs.tbl <- table(soil.samples[soil.samples])
obs.tbl

#Also tally soil types for each mule deer sampled
obs.tbl2 <- table(soils.locs[soils.locs])
obs.tbl2
```

10\. Converts the counts to proportions

```{r}
obs <- obs.tbl/sum(obs.tbl)
obs

obs2 <- obs.tbl2/sum(obs.tbl2)
obs2
```
